---
title: "Performance analysis of restaurants since the adoption of new food rating system by King County in January 2017"
author: "Sakshi Madan"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: slidy_presentation
---

```{r echo = FALSE}
# Course: 5210 Communicating Data
# Purpose: Presentation
# Date: December 13, 2019
# Author: Sakshi Madan
```

```{r setup, warning = FALSE, echo = FALSE, message = FALSE}
# Clear environment
rm(list = ls(all = TRUE))

# Clear environmet of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)
  lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""),
         detach, character.only = TRUE, unload = TRUE)

# Load Packages
library(tidyverse) 
library(here) 
library(magrittr) 
library(scales)
```

```{r importing_db, warning=FALSE, echo = FALSE, message = FALSE}
inspection_data <- read_csv(here('Sample_Food_Inspection_Data.csv'))
```

```{r adjusting, warning=FALSE, message=FALSE, echo=FALSE}
# Adjusting data types and names of the variables of required ones
inspection_data <- inspection_data %>% 
                            mutate(
                            Inspection_Date =  as.Date(`Inspection Date`, "%m/%d/%Y"),
                            City =  as.factor(toupper(City)), 
                            Inspection_Business_Name =  as.factor(`Inspection Business Name`), 
                            Inspection_Score =  as.numeric(`Inspection Score`), 
                            Inspection_Closed_Business =  as.factor(`Inspection Closed Business`), 
                            Violation_Type =  as.factor(`Violation Type`), 
                            Violation_Points =  as.numeric(`Violation Points`), 
                            Business_ID =  as.factor(Business_ID), 
                            Inspection_Serial_Num =  as.factor(Inspection_Serial_Num), 
                            )



inspection_data$`Inspection Date` <- NULL
inspection_data$`Inspection Business Name` <- NULL
inspection_data$`Inspection Score` <- NULL
inspection_data$`Inspection Closed Business` <- NULL
inspection_data$`Violation Type` <- NULL
inspection_data$`Violation Points` <- NULL

#Looking for rows where inspection serial number is NA and will drop those rows
inspection_data <- inspection_data %>% drop_na(Inspection_Serial_Num)

#Factor level 'NONE' for violation types
inspection_data$Violation_Type <- fct_explicit_na(inspection_data$Violation_Type, na_level = "NONE")

#We still had 54 rows in the original dataset where inspection scores are NA, creating a new table for them
na_score <- inspection_data[is.na(inspection_data$Inspection_Score),] %>% select(14,19)

#Drop rows with duplicate inspection serial number
na_score <- na_score[!duplicated(na_score$Inspection_Serial_Num),]

#Looking for those inspection serial numbers with NA inspection score in the original table and summing their violation points to find out inspection score
for (row in 1:nrow(na_score)) {
  df <- inspection_data[inspection_data$Inspection_Serial_Num == na_score[row, "Inspection_Serial_Num"]$Inspection_Serial_Num,]
  score <- sum(df$Violation_Points)
  na_score[row, "Inspection_Score"]$Inspection_Score <- score
}

#Merging both the tables na_score and inspection_data
inspection_data <- merge(inspection_data,na_score, by = c('Inspection_Serial_Num'), all.x = TRUE)
inspection_data$Inspection_Score.x[is.na(inspection_data$Inspection_Score.x)] <- inspection_data$Inspection_Score.y[is.na(inspection_data$Inspection_Score.x)]

na_score <- NULL
inspection_data$Inspection_Score <- inspection_data$Inspection_Score.x
inspection_data$Inspection_Score.x <- NULL
inspection_data$Inspection_Score.y <- NULL

# Looking for those rows where inspection score < 0 in original dataset
negative_values <- inspection_data %>% 
  filter(Inspection_Score < 0) %>% 
  select(1,22)

for (row in 1:nrow(negative_values)) {
  df <- inspection_data[inspection_data$Inspection_Serial_Num == negative_values[row, "Inspection_Serial_Num"],]
  score <- sum(df$Violation_Points)
  negative_values[row, "Inspection_Score"] <- score
}

#Merging both the tables negative_values and inspection_data
inspection_data <- merge(inspection_data,negative_values, by = c('Inspection_Serial_Num'), all.x = TRUE)
inspection_data$Inspection_Score.x[inspection_data$Inspection_Score.x < 0] <- inspection_data$Inspection_Score.y[inspection_data$Inspection_Score.x < 0]

negative_values <- NULL
inspection_data$Inspection_Score <- inspection_data$Inspection_Score.x
inspection_data$Inspection_Score.x <- NULL
inspection_data$Inspection_Score.y <- NULL

#Working on inspection date
inspection_data$Year <- format(inspection_data$Inspection_Date,"%Y")

inspection_data %<>% mutate(Year = case_when(
                            Year == "0006" ~ "2006",
                            Year == "0007" ~ "2007",
                            Year == "0008" ~ "2008",
                            Year == "0009" ~ "2009",
                            Year == "0010" ~ "2010",
                            Year == "0011" ~ "2011",
                            Year == "0012" ~ "2012",
                            Year == "0013" ~ "2013",
                            Year == "0014" ~ "2014",
                            Year == "0015" ~ "2015",
                            Year == "0016" ~ "2016",
                            Year == "0017" ~ "2017",
                            Year == "0018" ~ "2018",
                            Year == "0019" ~ "2019")
                            )

inspection_data$Inspection_Date <- NULL

#selecting required columns
inspection_data_filtered <-  inspection_data %>% select(c(1,6,9,10,14,17,18,19,21,22))


#saving un-duplicated rows as per inspection serial number
inspection_data_not_dup <- inspection_data_filtered[!duplicated(inspection_data_filtered$Inspection_Serial_Num),]

```


```{r echo=FALSE, warning=FALSE, message=FALSE}
#For violations graph

unique_business_inspection_year <- inspection_data_not_dup %>% 
  group_by(Year) %>% 
  summarise(count_businesses = n_distinct(Business_ID))

#no violation
none_count <- inspection_data_filtered %>% 
  group_by(Year, Business_ID) %>% 
  summarise(count = length(which(Violation_Type != "NONE"))) %>% 
  filter(count == 0) %>% 
  group_by(Year) %>% 
  summarise(none = n())

none_count$ratio <- none_count$none / unique_business_inspection_year$count_businesses

#red violation
red_count_total <- inspection_data_filtered %>% 
  group_by(Year, Business_ID) %>%
  summarise(count = length(which(Violation_Type == "RED"))) %>%
  filter(count != 0) 

red_count <- red_count_total %>%
  group_by(Year) %>%
  summarise(red = n())

red_count$ratio <- red_count$red/unique_business_inspection_year$count_businesses

#blue violation
blue_count_total <- inspection_data_filtered %>% 
  group_by(Year, Business_ID) %>% 
  summarise(count = length(which(Violation_Type == "BLUE"))) %>%
  filter(count != 0)

red_count_total <- red_count_total %>% select(-3) 
blue_count_total <- blue_count_total %>% select(-3)

intersect <- inner_join(red_count_total, blue_count_total)
blue_count <- anti_join(blue_count_total, intersect, by = c('Year','Business_ID'))

blue_count <- blue_count %>%
              group_by(Year) %>%
              summarise(blue = n())

blue_count$ratio <- blue_count$blue/unique_business_inspection_year$count_businesses

```

```{r echo=FALSE, warning = FALSE, message = FALSE}
#For violations in cities graph

city_unique_business_inspection_year <- inspection_data_not_dup %>% 
  filter(City %in% c("BELLEVUE","KENT","FEDERAL WAY","SEATTLE","RENTON")) %>% 
  group_by(Year, City) %>% 
  summarise(count_businesses = n_distinct(Business_ID))

#none violation
city_none_count <- inspection_data_filtered %>% 
  filter(City %in% c("BELLEVUE","KENT","FEDERAL WAY","SEATTLE","RENTON")) %>% 
  group_by(Year, City, Business_ID) %>% 
  summarise(count = length(which(Violation_Type != "NONE"))) %>% 
  filter(count == 0) %>% 
  group_by(Year, City) %>% 
  summarise(none = n())

merged_none <- merge(city_unique_business_inspection_year, city_none_count, by = c('Year','City'))
merged_none$rate <- merged_none$none/merged_none$count_businesses


#red violation
city_red_count_total <- inspection_data_filtered %>% 
  filter(City %in% c("BELLEVUE","KENT","FEDERAL WAY","SEATTLE","RENTON")) %>% 
  group_by(Year, City, Business_ID) %>% 
  summarise(count = length(which(Violation_Type == "RED"))) %>% 
  filter(count != 0) 

city_red_count <- city_red_count_total %>% 
  group_by(Year, City) %>% 
  summarise(red = n())

merged_red <- merge(city_unique_business_inspection_year, city_red_count, by = c('Year','City'))
merged_red$rate <- merged_red$red/merged_red$count_businesses

```


Data and Agenda  
===============================================

<div style="float: left; width: 95%;">

+ Data
<font size="5">
    - Food Inspection data from January 2006 - November 2019
    - We had 274782 observations for food inspections. There were multiple observations for inspections which had more than one violation
    - Performance declines with higher inspection score
    - Violation Types: Critical(Red) and Non-Critical(Blue)
    - Restaurants with even one critical(red) violation in a particular year has been categorized under restaurants with critical violation
    - We have chosen 5 major cities for comparison - Seattle, Bellevue, Kent, Renton, Federal Way because they have the maximum number of restaurants in King County

</font>

+ Agenda
<font size="5">
    - Effect on median inspection score in King County
    - Effect on median inspection score in 5 major cities
    - Effect on number of restaurants with no violations, critical violations and blue violations
    - Effect on number of restaurants with no violations in 5 major cities
    - Effect on number of restaurants with critical(red) violations in 5 major cities

</font>
</div>

Effect on median inspection score in King County
===============================================

<center>

<br>
```{r echo = FALSE, message = FALSE}
median_inspection_score <- inspection_data_not_dup %>%
  filter(Inspection_Score != 0) %>%
  group_by(Year) %>%
  summarise(score = median(Inspection_Score)) 

subtitle = paste0("After the introduction of the new rating system in 2017, there was a decline of ", round((median_inspection_score$score[11] - median_inspection_score$score[12])/median_inspection_score$score[11]*100,2), "% median inspection score")

median_inspection_score %>%
  ggplot(mapping = aes(y = score, x = Year, group = 1)) +
  geom_line() +
  geom_vline(xintercept = 9, linetype="dotted", color = "dark grey", size=1) +
  geom_vline(xintercept = 12, linetype="dotted", color = "dark grey", size=1) +
  scale_x_discrete(labels = c('2006':'2019')) +
  labs(title = "New rating system leading to better inspection scores",subtitle = subtitle , y = "Median Inspection Score", x ="") +
  annotate("text", x = 7.9, y = 14.6, label = "Food Borne \nIllness Outbreak", size = 3, color ="dark grey") +
  annotate("text", x = 12.8, y = 14.6, label = "New Rating \nSystem", size = 3, color = 
             "dark grey") +
  theme(plot.title = element_text(size = 11) ,plot.subtitle = element_text(size = 9), axis.title = element_text(size = 9), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(color = "black"), legend.title = element_blank())
```
</center>

Effect on median inspection score in 5 major cities 
==============================================
<center>

<br>
```{r echo = FALSE, message = FALSE}
median_inspection_score_cities <- inspection_data_not_dup %>% 
  filter(City %in% c("BELLEVUE","KENT","FEDERAL WAY","SEATTLE","RENTON")) %>% 
  filter(Inspection_Score != 0) %>% 
  group_by(Year, City) %>% 
  summarise(score = median(Inspection_Score))

subtitle = paste0("Median inspection score of Bellevue has been ", median_inspection_score_cities[66,3], " in 2018 and 2019 when the other cities were at ", median_inspection_score_cities[70,3])

  
 median_inspection_score_cities %>% 
  ggplot(mapping = aes(x = Year, y = score, group = City, color = City, alpha = City, size = City)) + 
  geom_line() +
  labs(title = "Median inspection scores of all the cities is now the same except Bellevue", subtitle = subtitle, y = "Median Inspection Score") +
  scale_alpha_manual(values = c(0.6,0.4,0.4,0.4,0.6)) +
  scale_size_manual(values = c(3,2,1,1,1)) +
  scale_color_manual(values = c("darkorange3","cadetblue","purple4","yellow4","grey25")) +
  theme(plot.title = element_text(size = 11), plot.subtitle = element_text(size = 9), axis.title = element_text(size = 9), legend.position = "bottom",panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(color = "black"), legend.title = element_blank()) 
```

</center>

Effect on number of restaurants with no violations, critical violations and non-critical violations
============================================================================================

<center>

<br>
```{r echo = FALSE, message = FALSE}
none_count <- none_count %>% select(1,2)
red_count <- red_count %>% select(1,2)
blue_count <- blue_count %>% select(1,2)

merged_violations <- merge(unique_business_inspection_year,none_count, by = c('Year'), all.x = TRUE)
merged_violations <- merge(merged_violations, red_count, by = c('Year'), all.x = TRUE)
merged_violations <- merge(merged_violations, blue_count, by = c('Year'), all.x = TRUE)
merged_violations <- merged_violations %>% gather(type, count, none:blue)
merged_violations$rate <- merged_violations$count/merged_violations$count_businesses
merged_violations$label <- paste0(sprintf("%.0f", merged_violations$rate*100), "%")

merged_violations %>%
        ggplot(mapping = aes(x = Year, y = rate, fill = factor(type, levels = c('blue','red','none')))) +
        geom_bar(position = position_stack(), stat = "identity", width = 0.7) +
        geom_text(aes(label = label), position = position_stack(vjust = 0.5), size = 3, color = 'black') +
        coord_flip() +
        scale_fill_manual(values = c('steelblue4','red4','snow4'), name = "Violation Type:", labels = c("Blue","Red","None")) +
        theme(legend.position = 'bottom', axis.title.y = element_blank(), plot.title = element_text(size = 11), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(color = "black"), axis.text.x = element_blank(), axis.ticks.x = element_blank(), legend.title = element_text(size = 8), legend.text = element_text(size = 8), legend.key.size = unit(0.8,"line"),legend.margin=margin(0,0,0,0),legend.box.margin=margin(-3,-3,-3,-3)) +
        labs(title = "New rating system is leading to healthier restaurants",y = "") +
        guides(fill = guide_legend(reverse = TRUE))
```

</center>

Effect on number of restaurants with no violations in 5 major cities
============================================================================================

<center>

<br>
```{r echo = FALSE, message = FALSE}
subtitle = paste0("In Seattle, restaurants with no violations increased to ",round(merged_none[60,5]*100), "%, ",round(merged_none[65,5]*100), "% and ",round(merged_none[70,5]*100), "% in 2017, 2018 and 2019, respectively") 

merged_none %>% 
  filter(City %in% c("BELLEVUE","KENT","FEDERAL WAY","SEATTLE","RENTON")) %>% 
  ggplot(mapping = aes(y = rate*100, x = Year, group = City, color = City, alpha = City, size = City)) +
  geom_line() +
  labs(title = "Seattle is the only city with consistent growing percentage of non-violating restaurants since the introduction \nof new system in January 2017", x = "", y = "Percentage of non-violating restaurants", subtitle = subtitle) +
  scale_alpha_manual(values = c(0.4,0.4,0.4,0.4,0.8)) +
  scale_size_manual(values = c(1,1,1,1,3)) +
  scale_color_manual(values = c("darkorange3","cadetblue","purple4","yellow4","grey25")) +
  scale_y_continuous(labels = function(x) paste0(x, '%')) +
  theme(plot.title = element_text(size = 11), plot.subtitle = element_text(size = 9), axis.title = element_text(size = 9), legend.position = "bottom", legend.title = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(color = "black")) 
```

</center>

Effect on number of restaurants with critical(red) violations in 5 major cities
============================================================================================

<center>

<br>
```{r echo = FALSE, message = FALSE}
subtitle = paste0("In Seattle, restaurants with red violations declined to ",round(merged_red[60,5]*100), "% , ",round(merged_red[65,5]*100), "%  and ",round(merged_red[70,5]*100), "% in 2017, 2018 and 2019, respectively") 

merged_red %>% 
  filter(City %in% c("BELLEVUE","KENT","FEDERAL WAY","SEATTLE","RENTON")) %>% 
  ggplot(mapping = aes(y = rate*100, x = Year, group = City, color = City, alpha = City, size = City)) +
  geom_line() +
  labs(title = "Seattle is the only city with consistent declining percentage of red violation restaurants since the \nintroduction of new system in January 2017", subtitle = subtitle, x = "", y = "Percentage of red violation restaurants") + scale_y_continuous(labels = function(x) paste0(x, '%')) +
  scale_alpha_manual(values = c(0.4,0.4,0.4,0.4,0.8)) +
  scale_size_manual(values = c(1,1,1,1,3)) +
  scale_color_manual(values = c("darkorange3","cadetblue","purple4","yellow4","grey25")) +
  #scale_y_continuous(labels = percent) +
  theme(plot.title = element_text(size = 11), plot.subtitle = element_text(size = 9), axis.title = element_text(size = 9), legend.position = "bottom", panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(color = "black"), legend.title = element_blank())
```

</center>

Summary
===============================================

<div style="float: left; width: 95%;">

<br>
<font size="5">

+ Performance of restaurants are improving in terms of inspection score

+ Each city is responding differently to rating system

+ Seattle is the most consistent one among the 5 cities

+ Recommendations, questions and suggestions

+ My contact information, email: smadan@seattleu.edu

</font>
</div>

